---
title: "p8106_hw2_yg2625"
author: "Yue Gu"
date: "March 20, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret) # only for plot
library(splines)
library(lasso2) # only for data
library(mgcv)
library(tidyverse)
library(ggplot2)
library(tidyverse)
library(patchwork)
library(boot)
```

# Load Data
```{r}
concrete = read.csv("./data/concrete.csv") %>% 
  janitor::clean_names()

# matrix of predictors 
x <- model.matrix(compressive_strength~.,concrete)[,-1]
# vector of response
y <- concrete$compressive_strength

```

# (a) Create scatter plots of response vs. predictors using the function featurePlot().

```{r}
theme1 = trellis.par.get()
theme1$plot.symbol$col = rgb(.2, .4, .2, .5)
theme1$plot.symbol$pch = 16
theme1$plot.line$col = rgb(.8, .1, .1, 1)
theme1$plot.line$lwd = 2
theme1$strip.background$col = rgb(.0, .2, .6, .2)
trellis.par.set(theme1)
featurePlot(x, y, plot = "scatter", labels = c("","Y"),
            type = c("p"), layout = c(4, 2))
```

Based on the plots, we could observe that there is no linear relationship between the predictors and compressive strength except for cement. 


# (b) Perform polynomial regression to predict compressive strength using water as the predictor. For 1<=d<=4, use cross-validation to select the optimal degree d for the polynomial. What degree was chosen, and how does this compare to the results of hypothesis testing using ANOVA? Make a plot of different polynomial fits to the data.

## CV selection of optimal degree d
```{r}
set.seed(2)
delta = rep(NA, 4)
for (i in 1:4) {
  fit = glm(compressive_strength~poly(water,i), data = concrete)
  delta[i] = cv.glm(concrete, fit, K = 10)$delta[1]
}
plot(1:4, delta, xlab = "Degree", ylab = "Test MSE", type = "l")
points(which.min(delta), delta[which.min(delta)], col = "red", pch = 19, cex = 2)
```
Based on the CV plot, we choose degree of 4 for the smallest test MSE for $1\leq d\leq 4$.

## Hypothesis testing using ANOVA
```{r}
fit1 <- lm(compressive_strength~water, data = concrete)
fit2 <- lm(compressive_strength~poly(water,2), data = concrete) 
fit3 <- lm(compressive_strength~poly(water,3), data = concrete)
fit4 <- lm(compressive_strength~poly(water,4), data = concrete) 
anova(fit1, fit2, fit3, fit4)
```

Based on the result, model 4 with degree d = 4  provides the smallest MSE and it's the optimal model, which matches our choice from CV shown above.

## Creating plots for different polynomial fits
```{r}
pred1 = predict(fit1)
pred2 = predict(fit2)
pred3 = predict(fit3)
pred4 = predict(fit4)

p = ggplot(concrete, aes(x = water, y = compressive_strength)) +
  geom_point(color = rgb(.2, .4, .2, .5))
p1 = p + 
  geom_line(aes(x = water, y = pred1), concrete, color = rgb(.8, .1, .1, 1)) +
  ggtitle("d = 1")
p2 = p + 
  geom_line(aes(x = water, y = pred2), concrete, color = rgb(.8, .1, .1, 1)) +
  ggtitle("d = 2")
p3 = p + 
  geom_line(aes(x = water, y = pred3), concrete, color = rgb(.8, .1, .1, 1)) +
  ggtitle("d = 3")
p4 = p + 
  geom_line(aes(x = water, y = pred4), concrete, color = rgb(.8, .1, .1, 1)) +
  ggtitle("d = 4")

p1 + p2 + p3 + p4
```

Based on the plots, d = 4 provides the optimal model fit for its closest trend for data points.

# (c) Fit a smoothing spline using water as the predictor for a range of degrees of freedom, as well as the degree of freedom obtained by generalized cross-validation, and plot the resulting fits. Describe the results obtained.

## fit smoothing spline using water for a range of df
```{r}
waterlims = range(concrete$water)
water.grid = seq(from = waterlims[1],to = waterlims[2])


fit.ss1 = smooth.spline(concrete$water, concrete$compressive_strength,df = 30)
fit.ss2 = smooth.spline(concrete$water, concrete$compressive_strength,df = 40)
fit.ss3 = smooth.spline(concrete$water, concrete$compressive_strength,df = 50)
fit.ss4 = smooth.spline(concrete$water, concrete$compressive_strength,df = 60)
fit.ss5 = smooth.spline(concrete$water, concrete$compressive_strength,df = 70)
fit.ss6 = smooth.spline(concrete$water, concrete$compressive_strength,df = 80)
fit.ss7 = smooth.spline(concrete$water, concrete$compressive_strength,df = 90)
fit.ss8 = smooth.spline(concrete$water, concrete$compressive_strength,df = 100)
fit.ss9 = smooth.spline(concrete$water, concrete$compressive_strength,df = 110)

pred1 = as.data.frame(predict(fit.ss1, x = water.grid))
pred2 = as.data.frame(predict(fit.ss2, x = water.grid))
pred3 = as.data.frame(predict(fit.ss3, x = water.grid))
pred4 = as.data.frame(predict(fit.ss4, x = water.grid))
pred5 = as.data.frame(predict(fit.ss5, x = water.grid))
pred6 = as.data.frame(predict(fit.ss6, x = water.grid))
pred7 = as.data.frame(predict(fit.ss7, x = water.grid))
pred8 = as.data.frame(predict(fit.ss8, x = water.grid))
pred9 = as.data.frame(predict(fit.ss9, x = water.grid))

q1 = p + 
  geom_line(aes(x = x, y = y), data = pred1,
            color = rgb(.8, .1, .1, 1)) +
  ggtitle("df = 30")
q2 = p + 
  geom_line(aes(x = x, y = y), data = pred2,
            color = rgb(.8, .1, .1, 1)) +
  ggtitle("df = 40")
q3 = p + 
  geom_line(aes(x = x, y = y), data = pred3,
            color = rgb(.8, .1, .1, 1)) +
  ggtitle("df = 50")
q4 = p + 
  geom_line(aes(x = x, y = y), data = pred4,
            color = rgb(.8, .1, .1, 1)) +
  ggtitle("df = 60")
q5 = p + 
  geom_line(aes(x = x, y = y), data = pred5,
            color = rgb(.8, .1, .1, 1)) +
  ggtitle("df = 70")
q6 = p + 
  geom_line(aes(x = x, y = y), data = pred6,
            color = rgb(.8, .1, .1, 1)) +
  ggtitle("df = 80")

q1 + q2 + q3 + q4 + q5 + q6
```

Based on the plots, with increasing degree of freedom from 30 to 80, the flexibility also increases for each model. 

## fit smoothing spline using df obtained by generalized CV
```{r}
fit.ss <- smooth.spline(concrete$water, concrete$compressive_strength)
fit.ss$df


pred.ss <- predict(fit.ss,
                   x = water.grid)

pred.ss.df <- data.frame(pred = pred.ss$y,
                         water = water.grid)

p +
geom_line(aes(x = water, y = pred), data = pred.ss.df,
          color = rgb(.8, .1, .1, 1)) + theme_bw()

```

The plot above shows the optimal df calculated by generalized CV.

# (d) Fit a GAM using all the predictors. Plot the results and explain your findings.

## Fit GAM
```{r}
gam_fit = gam(compressive_strength ~ s(cement) + s(blast_furnace_slag) + s(fly_ash) + s(water) + s(superplasticizer) + s(coarse_aggregate) + s(fine_aggregate) + s(age), data = concrete)

summary(gam_fit)
```

## Plot the result
```{r}
par(mfrow = c(2, 4))
plot(gam_fit)
```

Based on the plots, we observe that there is a linear relationship between compressive strength and cement, where we could furtherly build linear regression model. However, there is no clear linear relationship between compressive strength and other predictors.

